Project-II by Group MexicoCity

### 2014-12-13 Project 2/report/paper.tex

@@

After observing the histogram of the counts, as proposed by the course staff, we can see that despite the high maximal count, most of the data is concentrated in the low-count range (1-3500). Going out on a limb here we might consider any count quite outside of this interval to be an outlier, however things are a bit nuanced: if we start removing outliers from the data (which is already sparse) we run into the problem of deciding how far the point has to be from this interval to consider it as such. That decision is very difficult to make, especially considering that music listening - as any human-  behaviour cannot be expected to always follow mathematical rules. \textcolor{red}{What about poisson distribution for count?}

@@

Then we applied the log transformation to obtain a more Gaussian input as seen in Figure \ref{fig:loghist}. We then compute the RMSE based on the same principle as before and rescale the result with the exponential to obtain an RMSE of 4130, which is already much better.

\subsection{Collaborative-filtering}

\section{Summary}

>>>>

We cross-validated parameters for the NN: namely the dropout, $L_2$ weight and dimension (depth mostly). Following are the corresponding plots, respectively. We can see that a dropout value of 0.43 is actually the best one to choose (the one with minimal test error, red curve) This was done for 15 values of the dropout, $k=4$, and 15 different seeds. We then cross-validated the $L_2$ weight penalty parameter for NN and obtained an optimal TPR of 81.45\% for a weight penalty of $10^{-6}$ (while setting the dropout to a static value of 0.43). We used 15 seeds and 20 values of the $L_2$ weight, with $k = 4$.
\begin{figure}
\centering     %%% not \center
\subfigure[NN with various dropouts]{\label{fig:histC}\includegraphics[width=0.49\textwidth]{images/NNKCVplotdropout.png}}
\subfigure[NN with various L2 weight penalties]{\label{fig:loghist}\includegraphics[width=0.49\textwidth]{images/NNKCVplotL2weight.png}}
\caption{Test errors}
\end{figure}

@@

After observing the histogram of the counts, as proposed by the course staff, we can see that despite the high maximal count, most of the data is concentrated in the low-count range (1-3500). Going out on a limb here we might consider any count quite outside of this interval to be an outlier, however things are a bit nuanced: if we start removing outliers from the data (which is already sparse) we run into the problem of deciding how far the point has to be from this interval to consider it as such. That decision is very difficult to make, especially considering that music listening - as any human-  behaviour cannot be expected to always follow mathematical rules.

@@

Then we applied the log transformation to obtain a more Gaussian input as seen in Figure \ref{fig:loghist}. We then compute the RMSE based on the same principle as before and rescale the result with the exponential to obtain an RMSE of 4130 (8.3 in log scale), which is already much better.

\subsection{Collaborative-filtering}
Let $UA$ (users-artists) be equal to $Y_train$. The values inside the matrix are the number of time an user heard an artist. One of the principle features of the problem now is to say whether a user “likes” an artist of not, and ideally to know how much.

We first tried to normalize each row independently so that we will get values between 0 and 1.  Then we applied the following method : We created the matrix $P$, which is a diagonal matrix where each value in the diagonal is the total listening count for each user, (there are problems with users with total count $= 0$, however those don’t exist in our dataset). Hence, this matrix is invertible.

Using this inverse we can compute the similarity between users. At this point we don’t take into account the social graph, but it looks promising to refer later (if we can assume that friends are more likely to like the same artists). The similarity matrix X should contain 1 in the diagonal ( a user is maximally similar to himself). However this is not the case in our dataset : when we computed the similarity of the user using the formula $UA\cdot UA^T$ then we obtain the total number of listening in common with other users. That means to get a diagonal of $1$’s, the matrix $P$ should be defined differently : instead of having a diagonal matrix where each entry is  total listening count of user $i$, we should simply have the same values as $UA \cdot UA^T$ in the diagonal matrix $P$ and $0$ elsewhere. Then we will measure the cosine similarity between another user. Note that this is possible to do without normalizing the data, as we obtain the same results regardless.

One idea is to use the social graph along with the similarity matrix we obtained is to see if it is indeed the case that friends are similar users ( listeners), but not the other way around. Therefore, when we want to add a new users and predict his counts (taste in music), one possibility would be to use the information in the social graph and then use the similarity matrix to predict.

So finding the social relation between users requires finding the position of $1$s in the social graph. With these positions we can check what is the similarity between friends and compute the average similarity. We obtained experimentally an average of $0.2145$, which means their taste are not as correlated as we thought, putting the usefulness of the social graph into question again. We can use this number when we want to predict counts for a new user: we can look into the social graph to obtain the relationship between the new user and all existing users and then decide on their similarity between them. To predict counts we can sum the counts for the friend's user for a particular artist and divide it by the number of his friends.
\section{Summary}
